Equalizer: A Mature Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
Minor tweaks
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@13 db7c848b-7755-4df2-b763-8796797afa52
Filling in more gaps
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@14 db7c848b-7755-4df2-b763-8796797afa52

In \cite{EMP:09}, we presented the Equalizer parallel rendering framework. This
paper also summarized the work in parallel rendering up to 2009. In the
following we will present the related work published since then.

@@

A canvas can define a frustum, which will create default, planar sub-frusta for
all of its segments. A segment can also define a frustum, which overrides the
canvas frustum, e.g., for non-planar setups such as CAVEs or curved
screens. These frusta describe a physically-correct display setup for a Virtual
Reality installation. A canvas may have a software or hardware swap barrier,
which will synchronize the rendering of all contributing GPUs.

@@

@@

\subsection{Barriers, Queues and Object Maps}

@@

\subsection{Application-specific Scaling}

Gullivers world

@@

\subsection{Swap Synchronization and GPU affinity}\label{sec:swap}

@@

Control workstation in VR setups

>>>>

In \cite{EMP:09}, we presented the Equalizer parallel rendering framework. This
paper also summarized the work in parallel rendering up to 2009. In the
following we will present the related work published since then. An extensive
Programming and User Guide provides in-depth documentation on using and
programming Equalizer\cite{Eilemann:13}.

@@

A canvas can define a frustum, which will create default, planar sub-frusta for
all of its segments. A segment can also define a frustum, which overrides the
canvas frustum, e.g., for non-planar setups such as CAVEs or curved
screens. These frusta describe a physically-correct display setup for a Virtual
Reality installation. A canvas may have a software or hardware swap barrier,
which will synchronize the rendering of all contributing GPUs. The software
barrier executes a \textsf{glFinish} to ensure the GPU is ready to swap, a
Collage barrier (\sref{SEC_barrier}) to synchronize all segments, the swap
buffers call followed by a \textsf{glFlush} to ensure timely execution of the
swap command. The hardware swap barrier is implemented using the
\textsf{WGL\_NV\_swap\_group} or \textsf{GLX\_NV\_swap\_group} extensions.

@@

\subsection{Runtime Reconfiguration}\label{SEC_reconfig}

Switching a layout, as described above, or switching the stereo rendering mode,
may involve a different set of resources after the change, including the launch
and exit of render client processes. Equalizer solves this through a
reconfiguration step at the beginning of each rendering frame. Each resource
(channel, window, pipe, node) has an activation count, which is updated when the
layout or any other relevant rendering parameter is changed. When a resource is
found whose activation count does not match its current start/stopped state, the
resource is created or destroyed and \textsf{configInit} or \textsf{configExit}
are called accordingly. In the current implementation, a normal configuration
initialization or exit, as described in \cite{EMP:09}, uses the same code path
with all used resources transitioning to a running or stopped state,
accordingly. Since starting new resources typically requires object mapping and
associated data distribution, it is a costly operation. To improve performance
and robustness, Equalizer will flush the rendering pipeline whenever a runtime
reconfiguration is detected, before executing the change. This implementation
allows the late join or early exit, transparent to application developers.

@@

\subsection{Barriers, Queues and Object Maps}\label{SEC_barrier}

@@

\subsection{Model Unit}

This feature allows applications to specify a scaling factor between the model
and the real world, to allow exploration of macroscopic or microscopic worlds in
virtual reality. The unit is per view, allowing different scale factors within
the same configuration. It scales both the specified projection surface as well
as the eye position (and therefore separation) to achieve the necessary effect.

@@

Applications can switch each view between mono and stereo rendering at runtime,
and can run both monoscopic and stereoscopic views concurrently
(\fig{FIG_teaser} (b)). While this is trivial for active stereo rendering, it
does potentially involve the start and stop of resources and processes for
passive stereo or stereo-dependent task decompositions, as described in
\sref{SEC_reconfig}.

@@

The monitor allows to reuse the rendering on another channel, typically for
monitoring a larger setup on a control workstation. Output frames on the display
channels are connected to input frames on a single monitoring channel. The
monitor equalizer changes the scaling factor and offset between the output and
input, so that the monitor channel has the same, but typically downscaled view,
as the originating segments.
